\toolname: file-based localization of numerical perturbations in data analysis pipelines

### 2020-04-08 paper.tex
finished methods edits/suggestions

We capture temporary files by replacing every
process P by a wrapper that first calls P and then saves the produced
temporary files to a read-only directory. This process replacement is done by pre-pending
 to the \texttt{PATH} environment
variable a directory that contains a wrapper script named after the executable
called by P.

Files written by multiple processes are disambiguated using a similar technique. For a
 file F written by the processes in \textbf{P} = \{$P_{1}$, \ldots,
 $P_{n}$\}, we first check that processes in \textbf{P} do not
 write concurrently to F, which would violate our assumptions. Then, we
 replace every process $P_{i}$ by a \texttt{PATH}-based wrapper that first
 calls $P_{i}$ and then saves F to a read-only directory. In this way,
 successive versions of F are preserved for comparison. We finally
 update the provenance graph accordingly, so that all files in the graph
 have an in-degree of 1 (Figure~\ref{fig:provenance-dag}). This operation also makes the provenance graph
 acyclic, since we assumed that a process could only release a single version of a file.

@@

After capturing transient files in the first condition, we re-run the pipeline
step by step in the second one to label processes. The output files
produced by a process in both conditions are compared: if no differences are found, the process is marked as
reproducible; otherwise, the process is marked as non-reproducible, and the
output files produced in the first condition are copied to the second one, to ensure
that differences do not propagate further in the pipeline. Processes are
instrumented transparently, through a modification of the \texttt{PATH}
variable similar to the one described previously. By default, differences
in output files are identified by comparing file checksums. Other
comparison functions can also be defined for specific file types, for
instance to ignore file headers or file sections containing timestamps.
\toolname finally creates a labeled
provenance graph highlighting non-reproducible processes.

@@

We applied \toolname to the minimal
pre-processing pipelines released by the Human Connectome Project
(\href{https://www.humanconnectome.org}{HCP}), a leading initiative in
neurosciences.

@@

The HCP developed a set of pre-processing pipelines to process structural,
functional, and diffusion MRI data acquired in the project. We focus on HCP
pre-processing pipelines for structural data, and particularly
on PreFreeSurfer and FreeSurfer. 
A detailed description of the analyses done by these
pipelines is available in~\cite{glasser2013}. 
In summary, PreFreeSurfer consists of the following steps: 
\begin{itemize}
\item Gradient Distortion Correction (DC), 
\item Alignment and Anatomical Average (AAve), T1w(s), T2w(s),  
\item Anterior/Posterior Commissure Alignment (ACPC-A), 
\item Brain Extraction (BExt), 
\item \tristan{Ali, why isn't BBR cross-modal registration here and in the heatmap?}
\item Bias Field Correction (BFC), 
\item Atlas-Registration (AR).
\end{itemize}
And FreeSurfer consists of the following ones:
\begin{itemize}
\item Image downsampling, 
\item T1w image registration, 
\item T1w image segmentation,
\item Surface placement, 
\item Surface registration.
\end{itemize}
The average processing time per subject is approximately 2~hours for
PreFreeSurfer and 8~hours for FreeSurfer. The average output file size is
approximately 2.7~GB for PreFreeSurfer and 4.1~GB for FreeSurfer.
%\tristan{Ali, are these numbers per subject or for 20 subjects? it's an average of 20 subjects}
The \reprozip traced files show that approximately PreFreeSurfer opened  
83K files and created 7.7K processes, and FreeSurfer opened 62K files and 
created 4K processes in total for a single subject.

We randomly selected 20 unprocessed subjects 
from the HCP data release S500 
available in \href{https://db.humanconnectome.org}{the ConnectomDB repository}. 
For each subject, available data consisted of 1 or 2 T1 images, and 1 or 2 T2 images, with 
256 x 320 x 320 voxels of size 0.7 × 0.7 × 0.7 mm.
Acquisition protocols and parameters are detailed in~\cite{van2013wu}.

>>>>

We capture temporary files by replacing every
process $P$ by a wrapper that first calls $P$ and then saves the produced
temporary files to a read-only directory. This process replacement is done by pre-pending
 to the \texttt{PATH} environment
variable a directory that contains a wrapper script named after the executable
called by $P$.

Files written by multiple processes are disambiguated using a similar technique. For a
 file $F$ written by the processes in \textbf{P} = \{$P_{1}$, \ldots,
 $P_{n}$\}, we first check that processes in \textbf{P} do not
 write concurrently to $F$, which would violate our assumptions. Then, we
 replace every process $P_{i}$ by a \texttt{PATH}-based wrapper that first
 calls $P_{i}$ and then saves $F$ to a read-only directory. In this way,
 successive versions of $F$ are preserved for comparison. We finally
 update the provenance graph accordingly, so that all files in the graph
 have an in-degree of 1 (Figure~\ref{fig:provenance-dag}). This operation also makes the provenance graph
 acyclic, since we assumed that a process could only release a single version of a file.

@@

After capturing transient files in the first condition (i.e. operating system, library version, etc.),
we re-run the pipeline
step by step in the second one to label processes. The output files
created by a process in both conditions are compared: if no differences are found, the process is marked as
reproducible; otherwise, the process is marked as non-reproducible, and the
output files produced in the first condition are copied to the second one, to ensure
that differences do not propagate further in the pipeline. Processes are
instrumented transparently through a modification of the \texttt{PATH}
variable similar to the one described previously. By default, differences
in output files are identified by comparing file checksums. Other
comparison functions can also be defined for specific file types, for
instance to ignore file headers or file sections containing timestamps.
\toolname finally creates a labeled
provenance graph highlighting non-reproducible processes.

@@

We applied \toolname to the minimal
pre-processing pipelines released by the Human Connectome Project
(\href{https://www.humanconnectome.org}{HCP}), a leading initiative in
neuroimaging.

@@

The HCP developed a set of pre-processing pipelines to process structural,
functional, and diffusion MRI data acquired in the project. We focus on HCP
pre-processing pipelines for structural data, and particularly
on PreFreeSurfer and FreeSurfer. 
A detailed description of the analyses done by these
pipelines is available in~\cite{glasser2013}. 
In summary, the PreFreeSurfer pipeline consists of the following steps: 
\begin{itemize}
\item Gradient Distortion Correction (DC), 
\item Alignment and Anatomical Average (AAve), T1w(s), T2w(s),  
\item Anterior/Posterior Commissure Alignment (ACPC-A), 
\item Brain Extraction (BExt), 
\item \tristan{Ali, why isn't BBR cross-modal registration here and in the heatmap?}
\item Bias Field Correction (BFC), 
\item Atlas-Registration (AR).
\end{itemize}
And the FreeSurfer pipeline consists of the following:
\begin{itemize}
\item Image downsampling, 
\item T1w image registration, 
\item T1w image segmentation,
\item Surface placement, 
\item Surface registration.
\end{itemize}
The average processing time per subject is approximately 2~hours for
PreFreeSurfer and 8~hours for FreeSurfer. The average output file size is
approximately 2.7~GB for PreFreeSurfer and 4.1~GB for FreeSurfer.
%\tristan{Ali, are these numbers per subject or for 20 subjects? it's an average of 20 subjects}
The \reprozip traced files show that approximately PreFreeSurfer opened  
83K files and created 7.7K processes, and FreeSurfer opened 62K files and 
created 4K processes in total for a single execution, in this case corresponding to a single subject.

We randomly selected 20 unprocessed subjects 
from the HCP data release S500 
available in \href{https://db.humanconnectome.org}{the ConnectomDB repository}. 
For each subject, available data consisted of 1 or 2 T1w images and 1 or 2 T2w images, with 
$256 \times 320 \times 320$ voxels of size $0.7 \times 0.7 \times 0.7$ mm.
Acquisition protocols and parameters are detailed in~\cite{van2013wu}.
